---
title: "Portal Dataset Report"
output: html_document
---


```{r loadLibrariesData, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
surveys <- read.csv('../clean_data/portal_data_joined.csv')
```

# Introduction

On Feb 12th, 2022 our collaborator, Dr. Ron Wisconsin, sent me an initial dataset of kangaroo rats and other creatures that were recorded as part of the "Long-term monitoring and experimental manipulation of a Chihuahuan Desert ecosystem near Portal, Arizona" project.  The raw data files are stored in `raw_data` folder.

After some data cleaning, I created a csv of relatively clean data (`data/portal_data_joined.csv`) to read work with in R.
For more info on the data cleaning I performed please see `notes/2020-02-13-datacleaninglog.md`.
The dataset contains `r nrow(surveys)` observations of the following variables: `r colnames(surveys)`.
The observations begin in `r min(surveys$year)` and go until `r max(surveys$year)`.


## Preview of the data 
```{r previewData, echo=FALSE}
kable(head(surveys))
```

## Data Summaries

The data includes the following distribution of taxa.

```{r taxa, echo=FALSE}
surveys %>% count(taxa) %>% kable()
```

Genus counts by year, a subset.

```{r gemisByYear, echo=FALSE}
yearly_counts <- surveys %>%
  count(year, genus)

kable(head(yearly_counts))
```



```{r genusByYearGraph}
ggplot(data = yearly_counts, mapping = aes(x = year, y = n, color = genus)) +
    geom_line()
```












